---
title: "Template1"
author: "Studentnames and studentnumbers here"
date: "`r Sys.Date()`"
output: pdf_document
---

#heyyyyyyyy \# Set-up your environment

```{r package_install, include=FALSE}
install.packages("tidyverse")
install.packages("yaml")
install.packages("rmarkdown")
install.packages("cbsodataR")
install.packages("systemfonts")
install.packages("sf")
```

```{r packages, include=FALSE}
require(tidyverse)
require(rmarkdown)
require(systemfonts)
require(yaml)
library(dplyr)
library(stringr)
library(cbsodataR)
library(sf)
```

# Title Page

Include your names

Include the tutorial group number

Include your tutorial lecturer's name

# Part 1 - Identify a Social Problem

Use APA referencing throughout your document. [Here's a link to some explanation.](https://www.mendeley.com/guides/apa-citation-guide/)

## 1.1 Describe the Social Problem

Include the following:

- Why is this relevant?

- ...

# Part 2 - Data Sourcing

## 2.1 Load in the data

Preferably from a URL, but if not, make sure to download the data and store it in a shared location that you can load the data in from. Do not store the data in a folder you include in the Github repository!

```{r loading_data}

woningvoorraad <- read.csv("data/aantalwoningenNL.csv",)
huishouden_per_regio <- read.csv("data/huishoudens_samenstelling_regio.csv", sep=";")
vergunde_nieuwbouw <- read.csv("data/vergunde_nieuwbouw.csv", sep=";")
wv_inkomen <- read.csv("data/subgroep_inkomen.csv", sep=";")


#table(wv_inkomen$Huishoudenskenmerken)

wv_inkomen$Huishoudenskenmerken <-ifelse(wv_inkomen$Huishoudenskenmerken == "Inkomen onbekend", NA, wv_inkomen$Huishoudenskenmerken) 
woningvoorraad$Totale.woningvoorraad..aantal. <- str_remove_all(woningvoorraad$Totale.woningvoorraad..aantal.,"\\.")


saveRDS(woningvoorraad, "data/aantalwoningenNl.rds")
saveRDS(huishouden_per_regio, "data/huishoudens_samenstelling_regio.rds")
saveRDS(vergunde_nieuwbouw, "data/vergunde_nieuwbouw.rds")
saveRDS(wv_inkomen, "data/subgroep_inkomen.rds")


```

midwest is an example dataset included in the tidyverse package

## 2.2 Provide a short summary of the dataset(s)

```{r}

woningvoorraad <-readRDS("data/aantalwoningenNl.rds")
huishouden_per_regio <-readRDS("data/huishoudens_samenstelling_regio.rds")
vergunde_nieuwbouw <-readRDS("data/vergunde_nieuwbouw.rds")
wv_inkomen <-readRDS("data/subgroep_inkomen.rds")


```

## Linking datasets

The datasets were linked using the variables "Regio.s" and "Perioden". These variables are present in both datasets and identify the same region and year.

A left join was chosen because the household dataset was used as the primary dataset. Housing stock information was added for all matching regions and years

```{r}

woningvoorraad <- woningvoorraad %>%
  filter(Status.van.bewoning=="Totaal")

huishouden_woningvoorraad <- left_join(huishouden_per_regio, woningvoorraad, by = c("Regio.s", "Perioden") )

```

In this case we see 28 variables, but we miss some information on what units they are in. We also don't know anything about the year/moment in which this data has been captured.

``` r
inline_code = TRUE
```

These are things that are usually included in the metadata of the dataset. For your project, you need to provide us with the information from your metadata that we need to understand your dataset of choice.

## 2.3 Describe the type of variables included

Think of things like:

- Do the variables contain health information or SES information?

- Have they been measured by interviewing individuals or is the data coming from administrative sources?

*For the sake of this example, I will continue with the assignment...*

# Part 3 - Quantifying

## 3.1 Data cleaning

```{r data_cleaning}
#mean(dataset$percollege)
```

Please use a separate 'R block' of code for each type of cleaning. So, e.g. one for missing values, a new one for removing unnecessary variables etc.

## 3.2 Generate necessary variables

variab namen

```{r}

names(huishouden_per_regio)

names(vergunde_nieuwbouw)

names(woningvoorraad)

names(wv_inkomen)

names(huishouden_woningvoorraad)


```

Variabelen hernoemen

```{r}
# Rename
names(huishouden_per_regio)[3] <- "Regio's"
names(huishouden_per_regio)[4] <- "Totaal_particuliere_huishoudens"
names(huishouden_per_regio)[5] <- "Totaal_eenpersoons_huishoudens"
names(huishouden_per_regio)[6] <- "Totaal_meerpersoons_huishoudens"
names(huishouden_per_regio)[7] <- "Totaal_meerpersoons_huishoudens_niet_gehuwed"
names(huishouden_per_regio)[8] <- "Totaal_meerpersoons_huishoudens_gehuwed"
names(huishouden_per_regio)[9] <- "Totaal_1_ouder_huishoudens"
names(huishouden_per_regio)[10] <- "Totaal_2 persoons_huishoudens"
names(huishouden_per_regio)[11] <- "Totaal_3 persoons_huishoudens"
names(huishouden_per_regio)[12] <- "Totaal_4 persoons_huishoudens"
names(huishouden_per_regio)[13] <- "Totaal_5_persoons_huishoudens"

names(vergunde_nieuwbouw)[3] <- "Regio's"
names(vergunde_nieuwbouw)[4] <- "Aantal_vergunde_nieuwbouw"


names(woningvoorraad)[3] <- "Regio's"
names(woningvoorraad)[4] <- "Aantal_woningvoorraad"
names(woningvoorraad)[5] <- "Aantal_koopwoningen"
names(woningvoorraad)[6] <- "Aantal_huurwoningen"
names(woningvoorraad)[7] <- "Van_woningcorporaties"
names(woningvoorraad)[8] <- "Overige_huurders"
names(woningvoorraad)[9] <- "Eigendom_onbekend_aantal"


names(wv_inkomen)[5]<- "Regio's"
names(wv_inkomen)[6]<- "Woningen_in_voorraad"


names(huishouden_woningvoorraad)[3] <- "Regio's"
names(huishouden_woningvoorraad)[4] <- "Totaal_particuliere_huishoudens"
names(huishouden_woningvoorraad)[5] <- "Totaal_eenpersoons_huishoudens"
names(huishouden_woningvoorraad)[6] <- "Totaal_meerpersoons_huishoudens"
names(huishouden_woningvoorraad)[7] <- "Totaal_meerpersoons_huishoudens_niet_gehuwed"
names(huishouden_woningvoorraad)[8] <- "Totaal_meerpersoons_huishoudens_gehuwed"
names(huishouden_woningvoorraad)[9] <- "Totaal_1_ouder_huishoudens"
names(huishouden_woningvoorraad)[10] <- "Totaal_2_persoons_huishoudens"
names(huishouden_woningvoorraad)[11] <-"Totaal_3_persoons_huishoudens"
names(huishouden_woningvoorraad)[12] <- "Totaal_4_persoons_huishoudens"
names(huishouden_woningvoorraad)[13] <- "Totaal_5_persoons_huishoudens"
names(huishouden_woningvoorraad)[15] <- "Aantal_woningvoorraad"
names(huishouden_woningvoorraad)[16] <- "Aantal_koopwoningen"
names(huishouden_woningvoorraad)[17] <- "Aantal_huurwoningen"
names(huishouden_woningvoorraad)[18] <- "Van_woningcorporaties"
names(huishouden_woningvoorraad)[19] <- "Overige_huurders"
names(huishouden_woningvoorraad)[20] <- "Eigendom_onbekend_aantal"
```

Variable 1

```{r gen_var1}
#variabelen maken:
huishouden_woningvoorraad$Aantal_woningvoorraad <- as.numeric(huishouden_woningvoorraad$Aantal_woningvoorraad)
huishouden_woningvoorraad$Totaal_particuliere_huishoudens <- as.numeric(huishouden_woningvoorraad$Totaal_particuliere_huishoudens)

huishouden_woningvoorraad$woningen_per_huishouden <- huishouden_woningvoorraad$Aantal_woningvoorraad/huishouden_woningvoorraad$Totaal_particuliere_huishoudens
```

Variable 2

```{r gen_var2}

# als tweede variabelen gaan wij een percentuele verdeling van alle woningen door heel nederland. Dus dan moeten wij het totale aantal woningen delen door de aantal beschikbare woningen per regio. Zo krijgen wij dus te zien hoeveel % van het totale aantal staat in een provincie

#woningen_nederland <- woningvoorraad %>% select(`Regio's`, Perioden, Aantal_woningvoorraad)
#woningen_nederland <- woningen_nederland %>%
  #filter(`Regio's`=="Nederland")
#woningen_nederland <- woningen_nederland %>% filter(Perioden == "2025")



woningen_provincies <- woningvoorraad %>% select(`Regio's`, Perioden, Aantal_woningvoorraad)
woningen_provincies <- woningen_provincies %>%
  filter(`Regio's`!="Nederland")
woningen_provincies <- woningen_provincies %>% filter(Perioden == "2025")

woningen_provincies$NLtotaal <- 8274478

woningen_provincies$Aantal_woningvoorraad <- as.numeric(woningen_provincies$Aantal_woningvoorraad)
woningen_provincies$NLtotaal<- as.numeric(woningen_provincies$NLtotaal)
woningen_provincies$provincie_van_geheel_procent <- woningen_provincies$Aantal_woningvoorraad/woningen_provincies$NLtotaal*100


```

## 3.3 Visualize temporal variation

To analyze the housing shortage in the Netherlands, it is essential to examine the number of new houses approved for construction each year. This figure shows the number of permits granted for new construction between 2015 and 2025. Changes over time can mean whether the supply of housing is growing enough to meet demand.

The income dataset contained observations with the category "Inkomen onbekend" (unknown income).

Because these observations could not be assigned to an income group, they were recoded as missing values (NA). This resulted in 18 missing observations, which were excluded from analyses involving income groups.

```{r}

table(wv_inkomen$Huishoudenskenmerken, useNA = "ifany")

vergunde_nieuwbouw$Perioden <- as.numeric(vergunde_nieuwbouw$Perioden)
#View permitted new construction (number) over the years.
ggplot( subset(vergunde_nieuwbouw, !is.na(Aantal_vergunde_nieuwbouw)), aes(x = Perioden, y = Aantal_vergunde_nieuwbouw) ) + geom_line(color = "darkviolet", linewidth = 1) + geom_point(color = "black", size = 2) + theme_bw() + labs( title = "Licensed new construction over time", x = "Year", y = "Number of permitted new construction") + scale_x_continuous(breaks = vergunde_nieuwbouw$Perioden) + scale_y_continuous(breaks = seq(60000, 95000, by = 5000)) + coord_cartesian(ylim = c(60000, 95000))

```

The figure shows that the number of approved housing projects has been changing over time. While permitting has increased in some years, the variation suggests that housing prices have not increased at a constant rate. Given the current housing crisis in the Netherlands,these developments are relevant because new construction is necessary to increase the housing stock.

## 3.4 Visualize spatial variation

To analyze spatial variation in the Dutch housing market, the total housing stock was compared across provinces in 2025. A map was chosen because geographical differences are easier to identify visually than in a table or standard chart.

The national total was excluded because the aim is to compare provinces.

To ensure a seamless merge with the geographical shapefile from the CBS, the province names had to be cleaned. This involved removing the administrative suffix " (PV)" from the names and recoding "Friesland" to its official name "Fryslân". Additionally, the housing stock variable was converted to a numeric format to allow for correct color mapping.

Finally, the cleaned housing stock data were joined to the CBS province shapefile using a left join, allowing the housing stock to be visualised geographically.

```{r visualise_map}
woning_map <- woningvoorraad %>% filter( Perioden == 2025, `Regio's` != "Nederland" ) %>% mutate( provincie = gsub(" \\(PV\\)", "", `Regio's`), Aantal_woningvoorraad = as.numeric(Aantal_woningvoorraad) )

woning_map$provincie[ woning_map$provincie == "Friesland" ] <- "Fryslân"

provincies <- cbs_get_sf( region = "provincie", year = 2025)

kaart_data <- left_join( provincies, woning_map, by = c("statnaam" = "provincie"))

#woning_map[, c("provincie", "Aantal_woningvoorraad")]
#kaart_data[ is.na(kaart_data$Aantal_woningvoorraad), c("statnaam")]
#names(provincies)
#unique(provincies$statnaam)
#class(woning_map$Aantal_woningvoorraad)

#Begin plot code
ggplot(kaart_data) + geom_sf(aes(fill = Aantal_woningvoorraad)) + labs(title = "Housing Stock by Province in the Netherlands (2025)", fill = "Housing stock") + theme_minimal() 

```

The map shows significant regional differences in housing stock across the Netherlands. Zuid-Holland, Noord-Holland and Noord-Brabant have the largest housing stocks, reflecting their larger populations and higher levels of urbanisation. Smaller provinces such as Drenthe, Flevoland and Zeeland have considerably fewer dwellings.

These regional differences are relevant to the Dutch housing problem, as they highlight the concentration of housing supply in highly urbanised regions where housing demand is also strongest. Understanding these regional differences helps identify areas where housing shortages may be most severe and where additional housing development may be needed.

Here you provide a description of why the plot above is relevant to your specific social problem.

## 3.5 Visualize sub-population variation

To investigate variation across sub-populations, housing stock levels were compared across income groups. Households with unknown income were excluded from the analysis. A boxplot was used to visualise differences in the distribution of housing stock between income categories over time.

```{r visualise_subpopulations}

wv_sub <- wv_inkomen %>% filter( Woningtype == "Totaal", `Regio's` == "Nederland", !is.na(Huishoudenskenmerken))

ggplot( wv_sub, aes( x = Huishoudenskenmerken, y = Woningen_in_voorraad) ) + geom_boxplot(fill = "pink") + labs( title = "Distribution of Housing Stock by Income Group (2018–2023)", x = "Income Group", y = "Number of Dwellings") + theme_minimal()

```

The boxplot shows differences in housing stock across income groups. Comparing the distributions allows us to assess whether certain income groups consistently account for a larger share of the housing stock. This is relevant to the housing shortage because housing availability may not be equally distributed across socioeconomic groups.


Here you provide a description of why the plot above is relevant to your specific social problem.

## 3.6 Event analysis

Analyze the relationship between two variables.

```{r analysis}

```

Here you provide a description of why the plot above is relevant to your specific social problem.

# Part 4 - Discussion

## 4.1 Discuss your findings

# Part 5 - Reproducibility

## 5.1 Github repository link

Provide the link to your PUBLIC repository here: ...

## 5.2 Reference list

Use APA referencing throughout your document.
